---
title: "Week6-discussion"
author: "Bailey Jørgensen"
format: html
---


## Setup 

::: {.callout-note}
## Recall our original plot from [lecture 4.1](https://eds-240-data-viz.github.io/course-materials/lecture-slides/lecture4.1-amounts-slides.html#/title-slide){target="_blank"}
Our "first draft" of this plot began in week 4 during our amounts / rankings lecture. We'll focus on improving colors and fonts today.
:::

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)
library(showtext)
library(glue)
library(ggtext)

#......................import Google fonts.......................
# `name` is the name of the font as it appears in Google Fonts
# `family` is the user-specified id that you'll use to apply a font in your ggpplot
font_add_google(name = "Josefin Sans", family = "josefin")
font_add_google(name = "Sen", family = "sen")

#....................import Font Awesome fonts...................
font_add(family = "fa-brands",
         regular = here::here("fonts", "Font Awesome 6 Brands-Regular-400.otf"))
font_add(family = "fa-regular",
         regular = here::here("fonts", "Font Awesome 6 Free-Regular-400.otf")) 
font_add(family = "fa-solid",
         regular = here::here("fonts", "Font Awesome 6 Free-Solid-900.otf"))

#......enable {showtext} rendering for all newly opened GDs......
showtext_auto()
showtext.opts(dpi = 300)

#..........................import data...........................
# find import code at: https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-03-05#grab-the-clean-data-here
jobs <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-05/jobs_gender.csv")

```


## Wrangle data

See [lecture 4.1 slides](https://eds-240-data-viz.github.io/course-materials/lecture-slides/lecture4.1-amounts-slides.html#/title-slide){target="_blank"} (specifically, [this slide](https://eds-240-data-viz.github.io/course-materials/lecture-slides/lecture4.1-amounts-slides.html#/data-wrangling){target="_blank"} and [this slide](https://eds-240-data-viz.github.io/course-materials/lecture-slides/lecture4.1-amounts-slides.html#/dumbbell-plot-subset){target="_blank"}) as a reminder of where we left off.

```{r}
#| eval: true
#| echo: true
#| warning: false
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                wrangle data                              ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

jobs_clean <- jobs |> 
  
  # add col with % men in a given occupation (% females in a given occupation is already included) ----
  mutate(percent_male = 100 - percent_female) |>
         
  # rearrange columns ----
  relocate(year, major_category, minor_category, occupation,
          total_workers, workers_male, workers_female,
          percent_male, percent_female,
          total_earnings, total_earnings_male, total_earnings_female, 
          wage_percent_of_male) |> 
  
  # drop rows with missing earnings data ----
  drop_na(total_earnings_male, total_earnings_female) |> 
  
  # make occupation a factor (for reordering groups in our plot) ----
  mutate(occupation = as.factor(occupation)) |> 
 
  # classify jobs by percentage male or female (these will become facet labels in our dumbbell plot) ----
  mutate(group_label = case_when(
    percent_female >= 75 ~ "Occupations that are 75%+ female",
    percent_female >= 45 & percent_female <= 55 ~ "Occupations that are 45-55% female",
    percent_male >= 75 ~ "Occupations that are 75%+ male"
  )) 

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                              create subset df                            ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#....guarantee the same random samples each time we run code.....
set.seed(0)

#.........get 10 random jobs that are 75%+ female (2016).........
f75 <- jobs_clean |> 
  filter(year == 2016, group_label == "Occupations that are 75%+ female") |> 
  slice_sample(n = 10)

#..........get 10 random jobs that are 75%+ male (2016)..........
m75 <- jobs_clean |> 
  filter(year == 2016, group_label == "Occupations that are 75%+ male") |> 
  slice_sample(n = 10)

#........get 10 random jobs that are 45-55%+ female (2016).......
f50 <- jobs_clean |> 
  filter(year == 2016, group_label == "Occupations that are 45-55% female") |> 
  slice_sample(n = 10)

#.......combine dfs & relevel factors (for plotting order).......
subset_jobs <- rbind(f75, m75, f50) |> 
  mutate(group_label = fct_relevel(.f = group_label, 
                                   "Occupations that are 75%+ female", 
                                   "Occupations that are 45-55% female", 
                                   "Occupations that are 75%+ male"),
         occupation = fct_reorder(.f = occupation, .x = total_earnings))
```

## Recreate original plot 

See [lecture 4.1](https://eds-240-data-viz.github.io/course-materials/lecture-slides/lecture4.1-amounts-slides.html#/dumbbell-plot-create){target="_blank"} as a reminder.

```{r}
#| eval: true
#| echo: true
#| warning: false
#| out-width: "100%"
# recreate original plot ----
plot <- ggplot(subset_jobs) +
  
  # create dumbbells ----
  geom_linerange(aes(y = occupation,
                     xmin = total_earnings_female, xmax = total_earnings_male)) + 
  geom_point(aes(x = total_earnings_male, y = occupation), 
             color = "#CD93D8", 
             size = 2.5) +
  geom_point(aes(x = total_earnings_female, y = occupation), 
             color = "#6A1E99", 
             size = 2.5) +
  
  # facet wrap by group ----
  facet_wrap(~group_label, nrow = 3, scales = "free_y") + # "free_y" plots only the axis labels that exist in each group
  
  # axis breaks & $ labels ----
  scale_x_continuous(labels = scales::label_currency(scale = 0.001, suffix = "k"),
                     breaks = c(25000, 50000, 75000, 100000))

plot
```

## Create new palette

```{r}
#| eval: true
#| echo: true
#| warning: false
# create a named color palette ----
earnings_pal <- c("males" = "#2D7787",
                  "females" = "#FC6B4B",
                  "dark_text" = "#0C1509",
                  "light_text" = "#4E514D") 

# preview it -----
monochromeR::view_palette(earnings_pal)
```

## Update plot colors & fonts

```{r}
#| eval: true
#| echo: true
#| warning: false
#| out-width: "100%"
#.........................create caption.........................

#........................create subtitle.........................

#..........................updated plot..........................
# update plot colors using palette ----
plot <- ggplot(subset_jobs) +
  
  # create dumbbells ----
  geom_segment(aes(x = total_earnings_female, xend = total_earnings_male,
                 y = fct_reorder(occupation, total_earnings), yend = occupation)) +
  geom_point(aes(x = total_earnings_male, y = occupation),
             color = earnings_pal["males"], size = 3.25) + # update with colors from `earnings_pal`
  geom_point(aes(x = total_earnings_female, y = occupation),
             color = earnings_pal["females"], size = 3.25) + # update with colors from `earnings_pal`
  
  # facet wrap by group ----
  facet_wrap(~group_label, nrow = 3, scales = "free_y") + # "free_y" plots only the axis labels that exist in each group
  
  # axis breaks & $ labels ----
  scale_x_continuous(labels = scales::label_dollar(scale = 0.001, suffix = "k"),
                     breaks = c(25000, 50000, 75000, 100000, 125000))

plot
```

```{r}
#.........................create caption.........................
github_icon <- "&#xf09b"
github_username <- "samanthacsik"

caption <- glue::glue(
  "Data Source: TidyTuesday (March 5, 2019)<br>
  <span style='font-family:fa-brands;'>{github_icon};</span>
  {github_username}"
)

#........................create subtitle.........................
money_icon <- "&#xf3d1"

subtitle <- glue::glue("<span style='font-family:fa-regular;'>{money_icon};</span>
                       Median earnings of full-time 
                       <span style='color:#2D7787;'>**male**</span> versus 
                       <span style='color:#FC6B4B;'>**female**</span>
                       workers by occupation in 2016")

#..........................updated plot..........................
updated_plot <- plot +
  labs(title = "Males earn more than females across most occupations",
       subtitle = subtitle,
       caption = caption) +
  theme_minimal(base_size = 7) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(family = "josefin",
                              face = "bold",
                              size = rel(1.7)),
    plot.subtitle = ggtext::element_textbox(family = "sen",
                                            size = rel(1.1),
                                            color = earnings_pal["light_text"],
                                            # fill = "lightblue", # for debugging 
                                            width = unit(25, "cm"),
                                            margin = margin(t = 2, r = 0, b = 6, l = 0),
                                            padding = margin(t = 5, r = 0, b = 5, l = 0)),
    plot.caption = ggtext::element_textbox(family = "sen",
                                           face = "italic", # NOTE: this no longer applies since the typeface "sen" does not exist in an italic font style
                                           color = earnings_pal["light_text"],
                                           # fill = "lightblue", # for debugging
                                           halign = 1, 
                                           lineheight = 1.5,
                                           width = unit(25, "cm"),
                                           margin = margin(t = 15, r = 0, b = 0, l = 0),
                                           padding = margin(t = 10, r = 0, b = 10, l = 0)),
    strip.text.x = element_text(family = "josefin",
                                face = "bold",
                                size = rel(1.1),
                                hjust = 0),
    panel.spacing.y = unit(0.5, "cm"),
    axis.text = element_text(family = "sen",
                             size = rel(0.8),
                             color = earnings_pal["light_text"]),
    axis.title = element_blank()
  )

updated_plot
```


## Save plot as a PNG file

```{r}
#| eval: true
#| echo: true

# write plot to file (aka save as png) ----
ggsave(
  filename = here::here("week6", "images", "salary-plot-5x5.png"),
  plot = updated_plot,
  device = "png",
  width = 5,
  height = 5,
  unit = "in",
  dpi = 300
)
```
